---
title: "README"
output: 
  html_document:
    keep_md: true
---

## 1- Merging test and train data set

* Read the data from test, train and merge them into a variable called merge_x
* The files do not have header so header is FALSE

```{r results='hide', cache=TRUE}
test_x <- read.table("UCI HAR Dataset/test/X_test.txt", header=FALSE, 
                   as.is = TRUE)
train_x <- read.table("UCI HAR Dataset/train/X_train.txt", header=FALSE,
                    as.is = TRUE)

merge_x <- rbind(test_x, train_x)
```

## 4 - Appropriately labels the data set with descriptive variable names

* Read the feature names from features.txt and assign as column names

```{r results="hide", cache=TRUE}
column_names <- read.table("UCI HAR Dataset/features.txt", 
                           header = FALSE, as.is = TRUE, 
                           colClasses = c("NULL", "character"), 
                           col.names=c("NULL","names") )$names

colnames(merge_x) <- column_names
```

## 2 - Extracts only the measurements on the mean and standard deviation for each measurement

* Filter  only the columns whose names match mean or std 

```{r results="hide", cache=TRUE}
merge_x <- merge_x[,grep('mean|std', names(merge_x), ignore.case = TRUE)]
```

## 3 - Uses descriptive activity names to name the activities in the data set

* Read activity_id from test, train and merge them

```{r results="hide", cache=TRUE}
test_y <- read.table("UCI HAR Dataset/test/y_test.txt", header=FALSE,
                     as.is = TRUE, colClasses = c(rep("character",1)),
                     col.names=c("activity_id"))
train_y <- read.table("UCI HAR Dataset/train/y_train.txt", header=FALSE,
                      as.is = TRUE, colClasses = c(rep("character",1)),
                      col.names=c("activity_id"))

merge_y <- rbind(test_y, train_y)
```

* Extract the actual activity_labels then map it to the activity id data.
* Finally add a column to the original data as activity_name

```{r results="hide", cache=TRUE}
activity_labs <- read.table("UCI HAR Dataset/activity_labels.txt", header=FALSE,
                            as.is=TRUE, colClasses=c(rep("character",2)), 
                            col.names=c("activity_id", "activity_name"))

merge_y <- merge(merge_y, activity_labs)$activity_name

merge_x$activity_name <- merge_y
```

## 5 -From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject

* Read the subject data from test,train and then merge them.
* Add an additional column to the original dataset

```{r results="hide", cache=TRUE}
test_sub <- read.table("UCI HAR Dataset/test/subject_test.txt", header=FALSE,
                     as.is = TRUE, colClasses = c(rep("character",1)),
                     col.names=c("subject"))
train_sub <- read.table("UCI HAR Dataset/train/subject_train.txt", header=FALSE,
                      as.is = TRUE, colClasses = c(rep("character",1)),
                      col.names=c("subject"))

merge_sub <- rbind(test_sub, train_sub)

merge_x <- cbind(merge_x, merge_sub)
```

* calculate mean of all columns using lapply
* write output to a result.txt

```{r results="hide", cache=TRUE}


library(data.table)
DT <- data.table(merge_x)
df <- DT[, lapply(.SD,mean), by=c("subject", "activity_name")]

write.table(df, file = "result.txt", row.names=FALSE, col.names = FALSE)
```